Figures for Paper

### 2020-11-24 paper_figures.Rmd
Adding more of the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes")
```

@@

pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
  # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
    p_value = lmp(model)
    r_squared = round(summary(model)$adj.r.squared,3)
    pretty_plot = broom::tidy(model) %>% 
        mutate(p_star = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  p.value > 0.1 ~ "NS")) %>% 
        filter(p.value < p_value_cutoff) %>%
        mutate(term = fct_reorder(term, -abs(estimate))) %>%
        ggplot(aes(x = term, y = estimate,fill = estimate > 0,show.legend = F)) +
        geom_col() +
        coord_flip() +
        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}")) +
        geom_text(aes(x=term, y= -0.01, label=p_star)) 
}
```

# Summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1)

@@

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals")

@@

Now I'm going to test out using mixed models.

```{r}
library(lme4)
unc_mixed = lmer(log10_unc13a_cryptic_leaf_psi_plusone ~ log10_stmn_2_cryptic_psi_leaf_plusone * number_g_alleles + (1 | pathology), data = new_regress)
summary(unc_mixed)
unc_mixed2 = lmer(log10_unc13a_cryptic_leaf_psi_plusone ~ log10_stmn_2_cryptic_psi_leaf_plusone * number_g_alleles+ (1 | pathology) + (1 | tissue_clean), data = new_regress)
summary(unc_mixed2)

@@

This leads us to exploring the affects of UNC13A psi separately for tissues
with 'low' STMN2 and those with 'high' STMN2. This distinction is simply cutting
the distribution of STMN2 psi in half and assigning each tissue to the 'low' or 
'high' group as so
```{r}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
if (!require("pacman")) install.packages("pacman")
library(pacman)
p_load("data.table",
       "tidyverse",
       "ggplot2",
       "rcompanion",
       "lme4",
       "ggpubr",
       "forcats",
       "scales",
       "here",
       "ggthemes",
       "ggrepel")
```

@@

pretty_effect_plot = function(model, p_value_cutoff = 0.05, title = "Model 1"){
  # Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
    p_value = lmp(model)
    r_squared = round(summary(model)$adj.r.squared,3)
    pretty_plot = broom::tidy(model) %>% 
        mutate(p_star = case_when(p.value < 0.001 ~ "***",
                                  p.value < 0.01 ~ "**",
                                  p.value < 0.05 ~ "*",
                                  p.value < 0.1 ~ ".",
                                  p.value > 0.1 ~ "NS")) %>% 
        filter(p.value < p_value_cutoff) %>%
        mutate(term = fct_reorder(term, -abs(estimate))) %>%
        ggplot(aes(x = term, y = estimate,fill = estimate > 0,show.legend = F)) +
        geom_col() +
        coord_flip() +
        ggtitle(title,subtitle = glue::glue("Model p-value: {p_value} \n Adj. R-squared: {r_squared}")) +
        geom_text(aes(x=term, y= -0.01, label=p_star)) 
}
```
# iPSC splicing

Read in the splicing results returned by MAJIQ and make a volcano plot, only highlight 
genes of interest with a label.

```{r}
junc_colors = c("#648FFF","#fe6101")
ipsc_splicing = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/ipsc_splicing_results.csv")

ipsc_splicing %>% 
  mutate(junction_name = glue::glue("{gene_name} - {paste_into_igv_junction}")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = ifelse(log10_test_stat == 16 & deltaPSI > 0 , junction_name,"")) %>%
  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
                  min.segment.length = unit(0, 'lines')) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  guides(alpha = FALSE, size = FALSE) + 
  scale_fill_manual(values = junc_colors) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab("-Log 10 Test Statistic") + 
  theme(text = element_text(size = 18))
```

# UNC13A Cryptic PSI across TDP-43 KD experiments

```{r}
rel_rna_cryptic_amount = fread("/Users/annaleigh/Documents/GitHub/unc13a_cryptic_splicing/data/kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv")

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

# Paient summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi_leaf), n = 2))) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                      stmn2_psi_groups == 2 ~ "High STMN2",
                                      TRUE ~ "No STMN2"))

@@

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals")

@@

We see that detection rate by genotype varies.

```{r}
overall_fisher = clean_data_table %>%

  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  dplyr::select(participant_id,unc13a_detected,call) %>% 
  unique() %>% 
  group_by(call) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(unc13a_detected)) %>% 
  dplyr::select(call,n_sample,n_sample_detected) %>% 
  unique()

over_p = overall_fisher %>% 
  column_to_rownames('call') %>% 
  fisher.test() %>% 
  broom::tidy() %>% 
  .$p.value

overall_fisher %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {n_sample} )")) %>% 
  ggplot(aes(x = detection_name, y = detection_rate, fill = detection_name)) +
  geom_col(show.legend = F) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739"))

@@

Although this difference is not significant, with the Fisher's exact giving
a p-value of `r round(over_p,2)`.

```{r}
detection_table = clean_data_table %>% 
  mutate(unc13a_detected = unc13a_cryptic_leaf_psi > 0) %>% 
  group_by(call,unc13a_detected,stmn2_psi_groups) %>% 
  add_count(name = "genotype_detected") %>% 
  dplyr::select(call,unc13a_detected,genotype_detected,stmn2_psi_groups) %>% 
  unique() %>% 
  pivot_wider(names_from = "unc13a_detected",
              values_from = "genotype_detected") %>% 
  dplyr::rename(unc_not_detected = `FALSE`, unc_cryptic_detected = `TRUE`) %>%
  mutate(total_tissue = (unc_not_detected + unc_cryptic_detected)) %>% 
  mutate(detection_rate = unc_cryptic_detected / total_tissue) %>% 
  ungroup() %>% 
  mutate(stmn2_psi_groups = fct_relevel(stmn2_psi_groups, "No STMN2", after = 0)) %>% 
  mutate(detection_name = glue::glue("{call} \n ( {total_tissue} )"))

count <- 0
breaks_fun <- function(x) {
  count <<- count + 1L
  switch(
    count,
    detection_table$detection_name[1:3],
    detection_table$detection_name[4:6],
    detection_table$detection_name[7:9],
  )
}

detection_table %>% 
  ggplot(aes(x = 1, y = detection_rate,fill = call)) + 
  geom_col(show.legend = F,position = 'dodge2') + 
  facet_wrap(~stmn2_psi_groups) +
  scale_fill_manual(values = c("#88CCEE","#44AA99","#1B7739")) + 
  scale_y_continuous(lim = c(0,0.5),labels = scales::percent) + 
  ylab("Percent of Samples \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18))  
  # xlab("N Tissues") +
  # scale_x_continuous(labels = breaks_fun, breaks = c(0.75, 1, 1.25), limits = c(0, NA)) 
```
